Supporting Urban Search \& Rescue Mission Planning\\ through Visualization-Based Analysis

### 2014-06-10 old/vis14-usar/PacficVis Reviewer Rebuttal.txt

We would like to thank the anonymous reviewers from PacificVis for their valuable
feedback that helped make the current revision much better and concise. In the following
we discuss how we addressed most of the reviewer`s comments.
The authors

Justification of design choices 
We included a requirements analysis and introduced the concept of sense-making and
decision-making theory to justify the decisions we made w.r.t. our system.

 Details of the implementation
  Performance and computation cost
We describe the details of the implementation in greater detail in the extended section 6.
The performance of the rendering, as well as the cost of the precomputation, is stated and
explained in this section as well.

 More information on the user feedback
We added a paragraph drawing conclusions from the feedback we received from the experts
during our qualitative evaluation.

As an application paper, a requirement analysis should be important. I
  hope the requirements and the supported tasks can be listed explicitly in
  the paper.
We adhered to this comment and included a requirement analysis and state the requirements
to our system in the introduction.

The workflow comparison in Section 3 is very important, since it tells
  the benefit of having the proposed system. However, it is not clear to me
  what are the differences of these two workflows. Maybe a list summarizing
  the differences would help.
We created a workflow figure (Figure 3) describing the differences between the current
workflow and our proposed solution.

For the path computation in Section 4.3, I wonder how many POIs can
  exist. If there are multiple POIs, how to plan the path with A*
  algorithm?
We added a discussion on this topic in Section 5.3

From Figure 1(bc), it seems that the point cloud rendering does not
  contain the ceiling/roof of the building. On the one hand, I suppose that
  the ceiling/roof data should be in the point cloud data, and should be
  important for spatial awareness. One the other hand, if the ceiling/roof
  is rendered, then the inner structure of the building will be blocked.
  This is not good. Therefore, I'm wondering how is ceiling/roof treated.
  Besides, a building usually has multiple floors. In this case, the
  occulsion problem will be serious in point cloud rendering. However, this
  is not mentioned in the paper.
This misunderstanding was clarified in Section 5.5 by adding a new part explaining the use
of interactive clip planes by the IC.

In the last paragraph of Section 4.3, it says "some might not arrive
  at the POI and are discarded". I wonder how is it possible. Given a start
  point and a POI, all paths calculated from A* algorithm should start from
  the start point and arrive at the POI. I suppose the only possibility of
  "not arrive" is that the path does not exist.
This point was clarified in Section 5.3 by explaining when a path does not exist.

In the first paragraph of Section 7, it says "to effectively annotate
  a three dimensional rendering of a filtered point cloud". I wonder how
  the point cloud is filtered.
The occupancy-based filtering is explained in greater detail in Sections 5.1 and 6.

The paper "prepared an expressive video showing the functionalities of
  out software".  The video helps much better understanding of the whole
  work. More information of this video should also be discussed in the
  paper, which is critical for readers to understand. 
We followed this suggestion and incorporated the video`s information into the paper in
various places.

First, many of the images chosen seem to be quite a poor representation
  of the system features. The teaser images were good; however, Figure 3
  left much to be desired.  Still images from the video could have been
  used to much greater effect (the video was very well done!).  In fact,
  the workflow in the video shows lots of really cool features which are
  poorly described in the paper.  
Figure 3 has been removed form the paper and the focus was shifted away from the
comparison with point-based rendering. We included more of the information contained in
the video into the paper.

Next, technical issues are glossed over.  Specifically Section 3 in the
  current workflow and visualization workflow, there is no sense of the
  amount of time it takes to generate these layouts, the co-registration,
  the fly-through, etc.  This was voiced by the users in the informal
  evaluation.  The runtime of the path computation should also be provided.
   If the goal of this system is to enable first responders, calculation
  time needs to be provided for assessment.
The timings for precomputations, renderings. path computations were added to the
Implementation section 6. The runtime for the co-registration is negligible as it is
performed during the real-time exploration (this is explained alongside the SLAM
reference)

So, from the abstract, the goal of this paper was to provide users
  information in a way that is easy to comprehend, and to "describe the
  design considerations for our system..." It seems likely that the
  consideration of path distance, inclination of walking surface, etc.
  would actually fit directly into a risk equation and provide some sort of
  metric (of course drilling down into multidimensional data is good and
  using PCP or SPLOMs can help).  However, from the informal feedback, the
  idea of assessing risk was not really discussed.  Also, in designing the
  scatterplot to handle four variables, using color and size as distinct
  variables can result in issues as color perception is directly related to
  the size of the area which is being colored.
This misunderstanding was hopefully removed by a more thorough description of the metric,
as this is the risk equation the reviewer is correctly talking about. The scatterplot
was replaced by a scatterplot matrix and the variable color and size were removed.

More details on the datatype for the point clouds should be given.  What
  is the size of the data, what fields are in the data (earlier there was a
  discussion of thermal for bodies and sound for heartbeats, is this
  captured?)
This point is now discussed in the Implementation section 6.

To summarize, this work is extremely interesting and the authors have
  made a large amount of good headway.  However, the paper needs a clear
  organization and goal.  If the goal is design constraints, it should
  focus more on this, describe how you can interact with the system,
  provide more details that are captured in the video.  If the paper is
  more focused on rendering of the building, then different tactics need to
  be chosen.  There is a lot of time spent on the point cloud rendering,
  but the technique used doesn't produce exceptionally great results. 
  However, that's not the crux of the work. Instead the authors wind up
  glossing over issues and leaving our crucial details.  For example, at
  the end of Section 4.5.2, the authors state "In this section we will
  describe the implementation details necessary to reproduce this work." 
  Then there are two sentences which basically states to use the OpenMP
  framework.  Is that all it takes for this?
The focus of the paper was removed from the point cloud rendering and shifted towards the
decision-making support of the system. Furthermore, the requested implementation section
was improved and describes the details.

Related work:
  The authors should also look at the following papers:
  William Ribarsky, Kalpathi R. Subramanian, Jianfei Liu, Onyewuchi
  Obirieze, Jack Guest: Mobile application for first response and emergency
  evacuation in Urban settings. COM.Geo 2010

SungYe Kim, Ross Maciejewski, Karl Ostmo, Edward J. Delp, Timothy F.
  Collins, David S.Ebert. Mobile Analytics for Emergency Response and
  Training. The Journal of Information Visualization, 7(1): 77-88, March
  2008.
Both papers were very relevant and were added to the discussion of related work.

The axes of the plots in Figure 6 are unreadable.  With overlaps in the
  profile plot, it is difficult to see which ones end earlier than others.
  Perhaps other visual encodings could have been used, or highlighting to
  help see the shortest paths?
The axes of the plots were redone. Furthermore, the plots are presented bigger to improve
readability more.

- Add a system overview, with the various components. Also, you should
  describe the implementation. In principle, the paper needs to be as
  reproducible as possible, and right now, it is not possible to know
  what's "under the hood".
A workflow figure was added to the paper, as well as an overview screenshot of the
complete system. The implementation is described in more detail in Section 6.

### 2014-06-11 vmv14-usar.tex
vmv14-usar: finished draft
git-svn-id: https://scivis.itn.liu.se/svn/publications/workinprogress/vmv14-usar@5443 cd0c9ba3-fe29-0410-b0c2-9a6eeba77867

%\authorfooter{
%    \item Alexander Bock and Timo Ropinski are with the Scientific Visualization Group, Link\"oping University. e-mail: \{alexander.bock\textbar timo.ropinski\}@liu.se
%    \item Jonas Lundberg is with the Graphic Design Group, Link\"oping University. e-mail: jonas.lundberg@liu.se
%    \item Alexander Kleiner is with the Collaborative Robotics Group, Link\"oping University. e-mail: alexander.kleiner@liu.se
%}

%\shortauthortitle{Bock \etal: Urban Search \& Rescue Mission Planning}

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
%\keywords{Urban search~\&~rescue mission planning, decision support, visual analysis.}

@@

\teaser{
	\centering
	\subfigure[Voxelized 3D point cloud rendering enabling an increased awareness of a damaged office building.]{
		\fbox{\includegraphics[height=0.26\linewidth]{figures/image1.jpg}}\label{fig:teaser:1}
	}
	\hfill
	\subfigure[Viable paths through buildings of the Tohoku university with two hazardous environments highlighted.]{
		\fbox{\includegraphics[height=0.26\linewidth]{figures/image2.jpg}}\label{fig:teaser:2}
	}
	\hfill
	\subfigure[Top-down view of the area providing contextual information.]{
		\fbox{\includegraphics[height=0.26\linewidth]{figures/image2-overview.jpg}}\label{fig:teaser:3}
	}
	\hfill
	\subfigure[Analysis of a path ensemble using parallel coordinates plots and profile plots.]{
		\fbox{\includegraphics[height=0.26\linewidth]{figures/pcpprofile.png}}\label{fig:teaser:4}
	}
  \caption{Our proposed visualization system applied to path planning in a collapsed building at Tohoku university. Different views (a--c) present the operator with an understanding of the building, allowing him to select and inspect paths that reach a point of interest. The assessment of the trade-off between paths is supported by a set of interactive visual analysis tools (d).}
  \label{fig:teaser}
}

@@

Natural disasters causing structural damage to inhabited buildings are an ever-present danger. As time-to-rescue is one of the most important factors determining victims' survivability, it is of highest importance for rescue responders to reach victims as fast as possible in order to provide medical attention or extraction. Planning and executing paths through buildings, however, is difficult due to structural damages making available floor plans outdated. Furthermore, structural weaknesses as well as hazardous environments, make regions and areas inaccessible. While trained responders can use their knowledge and intuition to spot these hazards, it is paramount that the \emph{Incident Commander} (IC) can analyze the information and coordinate multiple rescue responders. In order to design a computer support system that amplifies human decision-making, it is vital to consider how decisions in dynamic situations are made~\cite{Lundberg2012}. The system is then designed to support planning based on experience, to increase quality of control, and to support decision-making, not only with regard to expected decisions, but also considering unexpected developments.

There exist well-defined protocols describing the actions taken during an \emph{Urban Search~\&~Rescue} (USAR) operation. The Federal Emergency Management Agency describes five steps that are performed by the rescue team~\cite{fema08}. During the assessment step, 2D maps of the collapsed building are drawn based on the descriptions of rescue responders moving about the building searching for victims. In this exploration phase, they might stumble into hazardous areas that endanger the rescuer's life.

In recent years, technological developments made it possible to use unmanned vehicles to perform the initial exploration. The robots can be deployed into the building and are equipped with sensors that can detect victims, gather information about potentially hazardous environments, and perform detailed scans of rooms to create a 3D point cloud of the interior. After the map has been acquired, the \IC\ can analyze the collected data and plan viable access paths entering the building to reach certain \emph{Points of Interest} (POIs). In most cases, these POIs are potential locations of victims, but they can also be other mission critical areas that need to be reached and analyzed.

In this paper, we propose a visualization system that creates an interactive 3D rendering tailored to increase the \IC 's awareness of internal structures (see Figure~\ref{fig:teaser}~(a-c)) and supporting the analysis and planning of access paths (see Figure~\ref{fig:teaser}~(d)). The \IC\ can annotate the visualization with real-time information from on-site rescue responders, thus shifting the opportunistic decision making to being strategical. Our system computes an ensemble of access paths from the entry point to the POIs, where each path is based on varying risk factors. Uncertainty in the data and a priori unknown information make it infeasible to employ automatic algorithms to detect the globally optimal path. Our system supports the \IC\ in the process of analyzing and comparing all available paths at once to minimize the rescuer's danger and travel time.

@@

\section{Related Work} \label{sec:relatedwork}
\noindent {\bfseries Emergency management.} Much of the visualization-oriented work published in the field of emergency management is concerned with evacuation planning well before any rescue operation has to be performed. Notable work was performed by Reddy~\etal, which enables analyzing possible bottlenecks of escape routes~\cite{EuroVA12:13-17:2012}. Ribarsky~\etal\ presented a system organizing first responders in intact structures~\cite{Ribarsky:2010}. Kim~\etal\ developed a system enhancing the situational awareness of responders using a mobile visual analytics tool~\cite{Kim:2008}.

Many existing planning systems deployed nowadays in USAR scenarios are based on 2D representations~\cite{kleiner_et_al_ssrr09,KohlbrecherMeyerStrykKlingaufFlexibleSlamSystem2011,Pellenz2009SMU}. Given the 2D map of the environment, one common approach to path planning is to plan the shortest trajectory and to follow this trajectory stepwise. In harsh environments, however, the safest path, not the shortest, may be desirable. Wirth~\etal\ introduced an exploration strategy and path planner that utilizes occupancy grid maps when planning to reach several targets at the same time~\cite{Wirth2007ETA1}. Preliminary extensions towards exploration in 3D with regard to detection of voids were introduced by Dornhege and Kleiner~\cite{dornhege2011frontier}.

\noindent {\bfseries Point cloud visualization.} As the availability of point cloud data increased in recent years, there has been much research on rendering techniques for this kind of data. Basic rendering capabilities are provided by the widely used Point Cloud Library~\cite{Rusu11ICRA}. Furthermore, work by Richter~\etal\ uses a level-of-detail structure to render massive point clouds at high frame rates~\cite{Richter:2010:ORV:1811158.1811178}. More recently, Pintus~\etal\ presented a rendering algorithm that enhances features of unstructured point clouds without preprocessing~\cite{Pintus:2011:RRM:2384495.2384513}.

@@

\section{Decision-Making Theory} \label{sec:theory}
To be able to design a decision support system for USAR missions, it is crucial to take knowledge about human decision making into account. Human decision makers tend to evaluate options serially in time-constrained situations. They attempt to find one viable plan that should work rather than attempting to generate and compare numerous plans in parallel. This theory has been described by Klein and Calderwood as \emph{Recognition Primed Decision-making} (RPD)~\cite{KleinCalderwood}. Initially, human experts look for similarities to previous situations with regard to relevant goals, things that were important to monitor, and possible actions. Then, they go through a process of mental simulation to consider whether the actions are applicable to the case at hand. They assess the ongoing situation, looking for violations and confirmations of expectancies, which may require reframing the situation. Klein and Calderwood suggest that ``displays and interfaces can be centered on decisions rather than around data flows''~\cite{KleinCalderwood}, emphasizing that systems can be built to enhance the workflow of mental simulation.

The \emph{Contextual Control Model} (COCOM) by Hollnagel and Woods describes how people rely on context when making decisions~\cite{hollnagel2005joint}. This theory recognizes that humans sometimes act with plans of lower quality, relying on the environment to make decisions opportunistically without having a  plan to reach the goal. The quality of their control of the situation can be described as scrambled, opportunistic, tactical, or strategic. The scrambled mode refers to decisions made without any idea of what to do. In the opportunistic mode, people rely on cues in the local context to decide on their next action. In tactical mode, they have an idea of how to achieve their goal before taking action---a plan. In strategic mode, the plan includes coordination with other simultaneous goals. The goal of our proposed system is to raise the quality of control from being opportunistic (as in the current workflow) to being strategic, thus enabling improved decision-making capabilities.

Turning to the granularity of plans, the \emph{Extended Control Model} (ECOM) describes plans in terms of a tactical level (setting goals), monitoring (making plans and overseeing plans), regulating (managing local resources), and tracking (performing and adjusting actions)~\cite{hollnagel2005joint}. This theory can be used to apprise what kind of planning support a system provides.  Moreover, it has been argued by Lundberg~\etal\ that it is important to support resiliency: ``Rather than merely selecting a response from a ready-made table, [the system] must adapt and create a suitable response; either by following ready-made plans for adaptation or by making sense of the situation and create responses during the unfolding event''~\cite{Lundberg2012}. Thus, in addition to supporting specific responses that may be foreseen, the system should also support working outside of prepared situations.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\columnwidth]{figures/workflow.pdf}
	\caption{A schematic timeline overview of the currently employed workflow (left) and our proposed system (right), showing all events (red) and actions (blue) split up into five distinct phases. Utilizing the additional actions (yellow) in parallel and thus enabling faster exploration, we decrease the overall time-to-rescue.}
	\label{fig:workflow:workflow}
\end{figure}

While there is no general definition of a rescue team in an USAR incident, in most protocols, for example in FEMA~\cite{fema08} or Emergency Management Australia~\cite{em35}, one \IC\ is responsible for a single building and instructs multiple rescue responders inside. In this section, we will describe the current workflow first (see Figure~\ref{fig:workflow:workflow}, left) and then propose a visualization-enhanced workflow supported by our system (see Figure~\ref{fig:workflow:workflow}, right). Roman numerals in the text refer to the corresponding phases in the figure.

\noindent {\bfseries Current workflow.} After arriving at the disaster scene (Phase \texttt{I}), the first step for the responders is to explore and secure the area outside the collapsed building (Phase \texttt{II}). No rescuer is allowed to enter the building before it is secured, which can take an hour or more to finish (Phase \texttt{III}). Then, the \IC\ determines entry points into the structure and directs rescuers to perform reconnaissance (Phase \texttt{IV}). Using constant radio communication, the rescuer inside the building slowly advances and reports his progress to the \IC, who draws a 2D map (see Figure~\ref{fig:workflow:sota}) (Phase \texttt{V}). This exposes the rescuer to unquantifiable risks, like gas leaks or dormant fires. This is clearly an example of opportunistic control, where decisions are made opportunistically based on feedback from the environment. Although responders may recognize situations, decisions regarding the path are limited to the extent of the exploration and to their view of the local environment. Global planning is limited further by the ability of responders to communicate relevant structural information accurately to the \IC, such as angles of turns, which cause hand-drawn maps to experience drift.

\noindent {\bfseries Visualization-enhanced workflow.} The initial phases \texttt{I} and \texttt{II} are the same as in the workflow above. However, while the responders secure the building, the most time-consuming of the initial tasks, unmanned robots are released into the structure and start measuring the inside of the structure~(see the parallel execution in phase \texttt{III}). The robots' sensors are able to detect many signs of victims using thermal cameras and heart beat detectors~\cite{6027084, Wu12Eulerian}, but as these measurements are uncertain, false positives and false negatives might occur. The same holds true for hazardous environments like fires, gas leaks, structurally unsafe areas, chemical spills, or radiation. The data retrieval and preprocessing are done in parallel with securing the perimeter, so that all information is available without delay when phase \texttt{IV} begins. Based on suggested POIs, the system computes optimal paths through the generated map. This reduces time-to-rescue (Phase \texttt{V}) as the rescuers do not need to explore the building to the same extent. The proposed system applies RPD to the whole situation and extends the number of available planning from local conditions to higher ECOM levels.

When planning the access paths, a variety of factors must be taken into account. The responder must maintain a safe distance from hazardous environments, avoid overhanging structures, and the ground must be both stable and level. As these variables are extracted from uncertain data, they are difficult to quantify and are subject to uncertainties. The \IC\ has to perform trade-offs to choose between alternatives, for example favoring a longer path that is faster to traverse over a more dangerous, shorter path. These requirements make the path selection a perfect example demonstrating how the human-in-the-loop paradigm benefits the decision-making.

While the \IC\ is instructing the rescuer to follow one path, the rescuer feeds back information about possible victims or new hazards. The \IC\ incorporates this information into the system. This is of high importance as features might not only have been missed by the robots, but detected features might change during the rescue operation. Fires can start or extinguish, subsequent structural collapses can make areas inaccessible, or debris is removed after the initial reconnaissance.

@@

Combining expertise in visualization, cognitive systems engineering, rescue robotics, and discussions with domain experts, we determined a set of requirements that must be fulfilled to make our system useful for the \IC. We employed theories from sense-making and decision-making (see Section~\ref{sec:theory}) to guide the design of our system to fully support the \IC. The visualization components are designed to comply with these theories and have been tested in a user study with external domain experts as described in Section~\ref{sec:evaluation}.

The components should fulfill the following requirements, which we have derived from our discussions with the collaborating experts:

\begin{description}
\item[R1] The system must increase spatial awareness by allowing for interactive exploration of the collapsed structure.
\item[R2] The system must enable the \IC\ to interactively annotate the acquired data to react to changing circumstances.
\item[R3] The \IC\ must be able to inspect multiple access paths and be able to compare them and make trade-offs.
\item[R4] The system must provide the tools to select a globally optimal path and allow for its execution.
\end{description}

The proposed system employs multiple linked views to address these requirements (see Figure~\ref{sec:overview:system}). In order to fulfill {\bfseries R1}, our system renders the point cloud interactively, preserving occlusion and helping the \IC\ to form a consistent model of the structure. The \IC\ can seamlessly annotate newly discovered entrances, hazards, POIs, and inaccessible areas, thus fulfilling {\bfseries R2}. To address {\bfseries R3} and {\bfseries R4}, we integrate a visual representation of the different paths into the 3D visualization, and provide separate in-depth analysis tools.

The following sections address the components of our system. Before the acquired data can be used, a preprocessing must be performed (Section~\ref{sec:overview:preprocessing}). Following the preprocessing, interactive data annotation is possible (Section~\ref{sec:overview:annotation}). We describe the path computation process together with the employed comparison metrics in Section~\ref{sec:overview:pathcomputation}. The analysis of the path ensemble is described in the same section, whereas Section~\ref{sec:overview:3dvisualization} provides details on the considerations that went into designing the 3D visualization components of our system.

\begin{figure}
    \centering
    \framebox[\columnwidth][c]{
        \includegraphics[width=\columnwidth]{figures/fig-overview-system2.png}
    }
    \caption{A screenshot showing our system for a typical scenario. The right view allows for detailed inspection and navigation. Each view can be maximized for in-depth inspection.}
    \label{sec:overview:system}
\end{figure}

\subsection{Data Preprocessing} \label{sec:overview:preprocessing}
The data preprocessing consists of two steps. First, the data is scan-converted to facilitate 3D visualization and path computation. Second, attributes coding spatial information are derived, which are later used during the path analysis.

>>>>

@@

\teaser{
	\centering
	\subfigure[Voxelized 3D point cloud rendering that enables an increased awareness of an office building.]{
		\fbox{\includegraphics[height=0.2375\linewidth]{figures/image1.jpg}}\label{fig:teaser:1}
	}
	\hfill
	\subfigure[Viable paths through buildings of the Tohoku university with two hazardous environments.]{
		\fbox{\includegraphics[height=0.2375\linewidth]{figures/image2.jpg}}\label{fig:teaser:2}
	}
	\hfill
	\subfigure[Top-down view of the area providing contextual information.]{
		\fbox{\includegraphics[height=0.2375\linewidth]{figures/image2-overview.jpg}}\label{fig:teaser:3}
	}
	\hfill
	\subfigure[Path analysis using parallel coordinates plots and profile plots.]{
		\fbox{\includegraphics[height=0.2375\linewidth]{figures/pcpprofile.png}}\label{fig:teaser:4}
	}
  \caption{Our proposed visualization system applied to a collapsed building at Tohoku university. Different views (a--c) present the operator with an understanding of the building, allowing him to select and inspect paths that reach a point of interest. The assessment of the trade-off between paths is supported by a set of interactive visual analysis tools (d).}
  \label{fig:teaser}
}

@@

Structural damage to inhabited buildings is an ever-present danger. As time-to-rescue is an important factors for victims' survivability, it is of highest importance for rescue responders to reach victims and provide medical attention or extraction. Planning and executing paths through buildings, however, is difficult as available floor plans are outdated. Furthermore, structural weaknesses as well as hazardous environments, make regions and areas inaccessible. While trained responders can use their knowledge and intuition to spot these hazards, it is paramount that the \emph{Incident Commander} (IC) can analyze the information and coordinate rescue responders. There exist well-defined protocols describing the actions taken during an \emph{Urban Search~\&~Rescue} (USAR) operation. In most protocols, one \IC\ is responsible for a single building and instructs multiple rescue responders inside. The Federal Emergency Management Agency describes steps that are performed by the rescue team~\cite{fema08}. During the assessment step, 2D maps of the collapsed building are drawn based on the descriptions of rescue responders moving about the building searching for victims. In this exploration phase, they might stumble into hazardous areas that endanger the rescuer's life.

In recent years, technological developments made it possible to use unmanned vehicles to perform the initial exploration. The robots are deployed and are equipped with sensors that can detect victims, gather information about potentially hazardous environments, and perform detailed scans of the interior. After the map has been acquired, the \IC\ can analyze the collected data and plan access paths that reach certain \emph{Points of Interest} (POIs). In most cases, these POIs are potential locations of victims, but they can also be other mission critical areas.

In this paper, we propose a visualization system that creates an interactive 3D rendering tailored to increase the \IC 's awareness of internal structures (see Figure~\ref{fig:teaser}~(a-c)) and supports the planning and analysis of access paths (see Figure~\ref{fig:teaser}~(d)). Our system computes an ensemble of access paths from the entry point to the POIs, where each path is based on varying weighting factors. Uncertainty in the data and a priori unknown information make it infeasible to employ automatic algorithms to detect the globally optimal path. Our system supports the \IC\ in the process of analyzing and comparing all available paths at once to minimize the rescuer's danger and travel time.

@@

\section{Related Work} \label{sec:relatedwork}
\noindent {\bfseries Emergency management.} Much of the visualization-oriented work  in the field of emergency management deals with evacuation planning. Notable work was performed by Reddy~\etal, which enables analyzing possible bottlenecks of escape routes~\cite{EuroVA12:13-17:2012}. Ribarsky~\etal\ presented a system organizing first responders in intact structures~\cite{Ribarsky:2010}. Kim~\etal\ developed a system enhancing the situational awareness of responders using a mobile visual analytics tool~\cite{Kim:2008}. Many existing planning systems deployed nowadays in USAR scenarios are based on 2D representations~\cite{kleiner_et_al_ssrr09,KohlbrecherMeyerStrykKlingaufFlexibleSlamSystem2011}. Given the 2D map of the environment, one common approach to path planning is to plan the shortest trajectory and to follow this trajectory stepwise. Wirth~\etal\ introduced an exploration strategy and path planner that utilizes occupancy grid maps when planning to reach several targets at the same time~\cite{Wirth2007ETA1}.

\noindent {\bfseries Point cloud visualization.} Basic rendering capabilities for point cloud data are provided by the widely used Point Cloud Library~\cite{Rusu11ICRA}. Furthermore, work by Richter~\etal\ uses a level-of-detail structure to render massive point clouds at high frame rates~\cite{Richter:2010:ORV:1811158.1811178}. More recently, Pintus~\etal\ presented a rendering algorithm that enhances features of unstructured point clouds without preprocessing~\cite{Pintus:2011:RRM:2384495.2384513}.

@@

\section{Decision-Making Theory} \label{sec:theory}
It is crucial to take knowledge about human decision making into account to design a decision support system. Decision makers tend to evaluate options serially in time-constrained situations; they attempt to find one viable plan rather than attempting to generate and compare numerous plans in parallel. This theory has been described by Klein and Calderwood as \emph{Recognition Primed Decision-making} (RPD)~\cite{KleinCalderwood}. Initially, experts look for similarities to previous situations with regard to relevant goals, things that were important to monitor, and possible actions. Then, they go through a process of mental simulation to consider whether the actions are applicable to the case at hand. They assess the ongoing situation, looking for violations and confirmations of expectancies, which may require reframing the situation. Klein and Calderwood suggest that ``displays and interfaces should be centered on decisions rather than around data flows''~\cite{KleinCalderwood}, emphasizing that systems can be built to enhance the workflow of mental simulation.

The \emph{Contextual Control Model} (COCOM) by Hollnagel and Woods describes how people rely on context when making decisions~\cite{hollnagel2005joint}. Humans sometimes act with plans of lower quality, relying on the environment to make decisions opportunistically. The quality of their control can be described as scrambled, opportunistic, tactical, or strategic. The scrambled mode refers to decisions made without any information. In the opportunistic mode, people rely on cues in the local context to decide on their next action. In tactical mode, they have an idea how to achieve their goal before taking action---a plan. In strategic mode, the plan includes coordination with other simultaneous goals. Our goal is to raise the quality of control from being opportunistic (as in the current workflow) to being strategic, thus enabling improved decision-making capabilities.

The \emph{Extended Control Model} (ECOM) describes plans in terms of a tactical level (setting goals), monitoring (making plans and overseeing plans), regulating (managing local resources), and tracking (performing and adjusting actions)~\cite{hollnagel2005joint}. This theory can be used to apprise what kind of planning support a system provides.  Moreover, it has been argued by Lundberg~\etal\ that it is important to support resiliency: ``Rather than merely selecting a response from a ready-made table, [the system] must adapt and create a suitable response; either by following ready-made plans for adaptation or by making sense of the situation and create responses during the unfolding event''~\cite{Lundberg2012}. Thus, in addition to supporting foreseeable responses, the system should also support working outside of prepared situations.

@@

\begin{figure}
	\centering
	\includegraphics[height=0.75\columnwidth]{figures/workflow.pdf}
	\caption{A schematic timeline overview of the currently employed workflow and our proposed system, showing events (red) and actions (blue) split up into five distinct phases. Utilizing the additional actions (yellow) in parallel enables faster exploration and decreased overall time-to-rescue.}
	\label{fig:workflow:workflow}
\end{figure}

In this section, we will describe the current workflow first (see Figure~\ref{fig:workflow:workflow}, left) and then propose a visualization-enhanced workflow supported by our system (see Figure~\ref{fig:workflow:workflow}, right). Roman numerals in the text refer to the corresponding phases in the figure.

\noindent {\bfseries Current workflow.} The first step for the responders is to explore and secure the area outside the collapsed building (Phase \texttt{I}). No rescuer is allowed to enter the building before it is secured, which can take an hour or more to finish (Phase \texttt{II}). Then, the \IC\ determines entry points and directs rescuers to perform reconnaissance (Phase \texttt{III}). The rescuer inside the building slowly advances and reports his progress to the \IC, who draws a 2D map (Phase \texttt{IV}). This exposes the rescuer to unknown risks, like gas leaks or dormant fires. This is a good example of opportunistic control, where decisions are made opportunistically based on feedback from the environment. Although responders may recognize situations, decisions regarding the path are limited to the extent of the exploration and their view of the local environment. Global planning is limited further by the ability of responders to communicate relevant structural information accurately to the \IC, such as angles of turns, which cause hand-drawn maps to experience drift.

\noindent {\bfseries Visualization-enhanced workflow.} While the responders secure the building, the most time-consuming of the initial tasks, unmanned robots are released into the structure and measure the inside of the structure~(parallel execution in phase \texttt{II}). The robots' sensors are able to detect victims using thermal cameras and heart beat sensors~\cite{6027084}, but as these measurements are uncertain, false positives and false negatives might occur. The same holds true for hazardous environments like fires, gas leaks, structurally unsafe areas, chemical spills, or radiation. Data retrieval and preprocessing are done in parallel with securing the perimeter, so that all information is available when phase \texttt{III} begins. Based on suggested POIs, the system computes optimal paths through the generated map, reducing time-to-rescue (Phase \texttt{IV}) as the rescuers do not need to explore the building to the same extent. The proposed system applies RPD to the situation and extends the available planning from local conditions to higher ECOM levels.

When planning the access paths, a variety of factors must be taken into account. The responder must maintain a safe distance from hazardous environments, avoid overhanging structures, and the ground must be stable and level. As these variables are extracted from uncertain data, they are subject to uncertainties. The \IC\ has to perform trade-offs to choose between alternatives, for example favoring a longer, faster path over a more dangerous, shorter path.

While the \IC\ is instructing the rescuer to follow one path, the rescuer feeds back information about victims or new hazards. The \IC\ incorporates this information into the system. This is of high importance as features might not only have been missed by the robots, but detected features might change during the rescue operation. Fires can start or extinguish, subsequent collapses can make areas inaccessible, or debris is removed after the initial reconnaissance.

@@

Combining expertise in visualization, cognitive systems engineering, rescue robotics, and discussions with rescue experts, we determined requirements to make our system useful for the \IC. The visualization components are designed to comply with theories on sense-making and decision-making and have been tested in a user study with external domain experts. The requirements are as follows:
\begin{description}
\item[R1] The system must increase spatial awareness by allowing for interactive exploration of the collapsed structure.
\item[R2] The system must enable the \IC\ to interactively annotate the acquired data to react to changing circumstances.
\item[R3] The \IC\ must be able to inspect multiple access paths, compare them, and make trade-offs.
\item[R4] The system must provide the tools to select a globally optimal path and allow for its execution.
\end{description}

The proposed system employs multiple linked views to address these requirements (see Figure~\ref{sec:overview:system}). In order to fulfill {\bfseries R1}, our system renders the point cloud interactively, preserving occlusion and helping the \IC\ to form a consistent model of the structure. The \IC\ can seamlessly annotate newly discovered entrances, hazards, POIs, and inaccessible areas, thus fulfilling {\bfseries R2}. To address {\bfseries R3} and {\bfseries R4}, we integrate a visual representation of the different paths into the 3D visualization, and provide specific in-depth analysis tools.

The following sections address the components of our system. Before the acquired data can be used, a preprocessing must be performed (Section~\ref{sec:overview:preprocessing}). We describe the path computation process and the analysis together with comparison metrics in Section~\ref{sec:overview:pathcomputation}. Section~\ref{sec:overview:annotation} explains the annotation of the point cloud. Section~\ref{sec:overview:3dvisualization} provides details on the considerations that went into designing the 3D visualization components of our system.

\begin{figure}
    \centering
    \framebox[0.9\columnwidth][c]{
        \includegraphics[width=0.9\columnwidth]{figures/fig-overview-system2.png}
    }
    \caption{A screenshot of our system for a typical scenario. The right view allows for detailed inspection and navigation.}
    \label{sec:overview:system}
\end{figure}

\subsection{Data Preprocessing} \label{sec:overview:preprocessing}
The retrieved data from the unmanned robots is an unstructured point cloud. One issue with rendering such point clouds is missing occlusion information and the non-uniform distribution of measured points. To avoid this problem, we perform a binning to obtain a three-dimensional voxel data structure on a regular grid. The voxel size is dependent on the scan resolution of the robot, and is a trade-off between resolving smaller details and increasingly noisy data. In our cases, voxel sizes of about 5\,cm were sufficient with regard to this trade-off. From this point on we call a measurement in the original point cloud a \emph{point} and refer to a position in the grid-based, binned point cloud as a \emph{voxel}.

\noindent {\bfseries Attribute derivation.} In the second part, derived attributes are computed that are used to determine the set of rescue paths and to support their analysis. We compute a \emph{hazard distance field} (Figure~\ref{fig:overview:precomputation}) that denotes the distance to the closest hazard points. The \emph{support field} (Figure~\ref{fig:overview:precomputation}) shows the available supporting area. This value determines whether there is enough floor available for a responder to walk on. The \emph{occupancy field} (Figure~\ref{fig:overview:precomputation}) denotes the number of points each voxel is based on. A higher occupancy means that the voxel contains more points in the original point cloud data and thus provides a higher certainty. The \emph{size field} shows for each voxel if a rescuer can fit into the space. We calculate two size values, one with the rescuer standing up, and a second while crouching. In order to determine the steepness of surfaces, we compute the least-squares fitted plane based on all the points in the original point cloud which are covered by a voxel. The normal of the resulting plane is used as the normal for the voxel.

\subsection{Data Annotation} \label{sec:overview:annotation}
Each voxel can belong to one of five classes. \emph{Unclassified} is the default type for all voxels. \emph{Start} voxels are accessible entry points. This means that paths can start from any of these points. \emph{POI} voxels are destination points for paths. These indicates that there is a potential victim or another mission-critical element at this location. \emph{Hazard} voxels have been declared as dangerous. Each hazard area has a normalized severity denoting how dangerous it is. \emph{Forbidden} voxels can only be declared by the \IC\ and are areas completely out of reach. These points are used when, for example, a corridor collapses and is not accessible anymore. The classification for each voxel is modified by directly interacting with the 3D rendering, thus fulfilling {\bfseries R2}.

\subsection{Path Computation} \label{sec:overview:pathcomputation}
We employ the A* algorithm for the path computations~\cite{4082128}. The algorithm works as follows: For the current voxel $x$, the estimated remaining distance is calculated for all unvisited neighboring voxels $y_i$. This value is the sum of the cost to reach $x$, the cost to move from $x$ to $y_i$, and the estimated cost to reach the target from $y_i$. From these voxels the lowest cost is chosen as the next candidate. For a comprehensive description of the algorithm we refer the reader to the book by Russel and Norvig~\cite{AStar}.
